A survey of generic IND-CCA2 transformations

### 2024-07-08 faster-fo-transform/cca-transform-survey/main.tex
Explained U-bot KEM security

    \begin{enumerate}
        \item $A$ has no access to $\pco$
        \item $A$ has no access to $\cvo$
        \item The challenge ciphertext $A$ receives is obtained using a truly random coin, but the challenge ciphertext $B$ expects is obtained using a pseudorandom coin $r \leftarrow G(m)$
    \end{enumerate}
\end{proof}

>>>>

    \begin{enumerate}
        \item $A$ has no access to $\pco$
        \item $A$ has no access to $\cvo$
        \item The challenge ciphertext $A$ receives is obtained using a truly random coin, but the challenge ciphertext $B$ expects is obtained using a pseudorandom coin $r \leftarrow G(m)$
    \end{enumerate}

    % TODO: finish writing this section
\end{proof}

\subsection{From PKE to KEM}
The second part of the modular FO transform (denoted the $U$ transform) takes a $\pke$ and outputs a $\kem$. Depending on whether the input $\pke$ is rigid and whether the output $\kem$ rejects invalid ciphertext implicitly or explicitly, the security requirements for the input $\pke$ and the construction of the $\kem$ will vary. \cite{hofheinz2017modular} presents four variations of the $U$ transform, which are listed in table \ref{tbl:u-transform-summary}

\begin{table}[H]
    \center
    \begin{tabular}{|c|c|c|c|}
        \hline
        Name & $\pke$ requirements & rejection & shared secret\\
        \hline
        $U^\bot$ & \monospace{OW-PCVA} & $\bot$ & $H(m, c)$ \\
        \hline
        $U^{\not\bot}$ & \monospace{OW-PCA} & $H(s, c)$ & $H(m, c)$ \\
        \hline
        $U^\bot_m$ & rigidity + \monospace{OW-VA} & $\bot$ & $H(m)$ \\
        \hline
        $U^{\not\bot}_m$ & rigidity + \monospace{OW-CPA} & $H(s, c)$ & $H(m)$ \\
        \hline
    \end{tabular}
    \caption{A summary of variants of $U$ transformations}\label{tbl:u-transform-summary}
\end{table}

For the remaining of this section we will present the four transformations and sketch proofs of their securities.

\subsubsection{KEM with explicit rejection from randomized PKE}
Let $\pke = (\keygen, \encrypt, \decrypt)$ be a public-key encryption scheme, and $H: \mathcal{M}_\pke \times \mathcal{C}_\pke \rightarrow \mathcal{K}_\kem$ be a hash function. The $U^\bot$ transformation outputs a $\kem^\bot = (\keygen^\bot, \encap^\bot, \decap^\bot)$, where the key generation routine remains unchanged: $\keygen^\bot = \keygen$. The encapsulation and decapsulation routines are described in figure \ref{fig:u-bot-routines}

\begin{figure}[H]
    \begin{minipage}{0.49\textwidth}
        \begin{algorithm}[H]
            \caption{$\encap^\bot(\pk)$}\label{alg:u-bot-encap}
            \begin{algorithmic}[1]
                \State $m \leftsample \mathcal{M}_\pke$
                \State $c \leftsample \encrypt(\pk, m)$
                \State $K \leftarrow H(m, c)$
                \State \Return $(c, K)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \hfill
    \begin{minipage}{0.49\textwidth}
        \begin{algorithm}[H]
            \caption{$\decap^\bot(\sk, c)$}\label{alg:u-bot-decap}
            \begin{algorithmic}[1]
                \State $\hat{m} \leftarrow \decrypt(\sk, c)$
                \If{$\hat{m} \in \mathcal{M}_\pke$}
                    \State $K \leftarrow H(\hat{m}, c)$
                    \State \Return $K$
                \EndIf
                \State \Return $\bot$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \caption{$U^\bot$ routines}\label{fig:u-bot-routines}
\end{figure}

Under the random oracle model, the security of $\kem^\bot$ depends on the security of the input $\pke$.

\begin{theorem}\label{thm:owpcva-implies-indcca2}
    For every \monospace{IND-CCA2} adversary $B$ against $\kem^\bot$, there exists an \monospace{OW-PCVA} adversary $A$ against the underlying $\pke$ such that

    \begin{equation*}
        \monospace{Adv}(B) \leq \monospace{Adv}(A)
    \end{equation*}
\end{theorem}

\begin{proof}
    For a sketch of proof, we argue that under the random oracle model, $H(m, c)$ is indistinguishable from a uniformly random sample from $\mathcal{K}_\kem$ from the adversary $B$'s perspective unless $B$ somehow queries $H$ on $(m, c)$. This means that an \monospace{OW-PCVA} adversary can perfectly simulate the decapsulation oracle using uniformly random samples, as long as the outputs from the simulated decapsulation oracle are consistent with the outputs from the hash functions.

    The only flaw in the simulated decapsulated oracle lies in the case when $B$ queries $H(m^\ast, c^\ast)$, but because $A$ has access to $\pco$, $A$ can detect that $B$ has successfully recovered $m^\ast$ and uses the recovered $m^\ast$ to win the \monospace{OW-PCVA} game. In other words, $A$ keeps running $B$ in a simulation until $B$ makes the special hash query, at which time $A$ simply terminates $B$ and outputs the answer.

    To make the simulated decapsulation oracle and the hash oracle consistent, $A$ maintains two tapes $\mathcal{L}^\decap$ and $\mathcal{L}^H$ that record the queries made to the decapsulation oracle and the hash oracles respectively:

    $$
    \begin{aligned}
        (c, K) \in \mathcal{L}^\decap &\Leftrightarrow K = \mathcal{O}^\decap(c) \\
        (m, c, K) \in \mathcal{L}^H &\Leftrightarrow K = H(m, c)
    \end{aligned}
    $$

    If $\mathcal{O}^\decap$ has been queried with some valid ciphertext $c$, then when $H$ is queried with a corresponding input $(m, c)$ where $m$ is the decryption of $c$, $H$ needs to output the same value as $\mathcal{O}^\decap$. Similarly, if $H$ has been queried with some valid plaintext-ciphertext pair $(m, c)$, then when $\mathcal{O}^\decap$ is queried with $c$, the two oracles need to output the same value. The details of the simulation are described in figure \ref{fig:u-bot-patched-oracles}

    \begin{figure}[H]
        \begin{minipage}{0.49\textwidth}
            \begin{algorithm}[H]
                \caption{$\mathcal{O}^\decap_1(c)$}\label{alg:u-bot-patched-decap}
                \begin{algorithmic}[1]
                    \If{$\exists (\tilde{c}, \tilde{K}) \in \mathcal{L}^\decap : \tilde{c} = c$}
                        \State \Return $\tilde{K}$
                    \EndIf
                    \If{$\cvo(c) = 1$}
                        \State $K \leftsample \mathcal{K}_\kem$
                        \State Append $(c, K)$ to $\mathcal{L}^\decap$
                        \State \Return $K$
                    \EndIf
                    \State \Return $\bot$
                \end{algorithmic}
            \end{algorithm}
        \end{minipage}
        \hfill
        \begin{minipage}{0.49\textwidth}
            \begin{algorithm}[H]
                \caption{$H_1(m, c)$}\label{alg:u-bot-patched-hash}
                \begin{algorithmic}[1]
                    \If{$\exists (\tilde{m}, \tilde{c}, \tilde{K}) \in \mathcal{L}^H:(\tilde{m}, \tilde{c}) = (m, c)$}
                        \State \Return $\tilde{K}$
                    \EndIf
                    \State $K \leftsample \mathcal{K}_\kem$
                    \If{$\pco(m, c) = 1$}
                        \If{$\exists (\tilde{c}, \tilde{K}) \in \mathcal{L}^\decap : \tilde(c) = c$}
                            \State \Return $\tilde{K}$
                        \Else
                            \State Append $(c, K)$ to $\mathcal{L}^\decap$
                        \EndIf
                    \EndIf
                    \State Append $(m, c, K)$ to $\mathcal{L}^H$
                    \State \Return $K$
                \end{algorithmic}
            \end{algorithm}
        \end{minipage}
        \caption{Patched oracles in $U^\bot$}\label{fig:u-bot-patched-oracles}
    \end{figure}
\end{proof}

\subsubsection{KEM with implicit rejection from randomized PKE}
Given $\pke = (\keygen, \encrypt, \decrypt)$, the $U^{\not\bot}$ transformation outputs $\kem = (\keygen^{\not\bot}, \encap^{\not\bot}, \decap^{\not\bot})$. With implicit rejection, decapsulation routine returns some output from $H$ even when the input ciphertext $c$ is malformed. Specifically, the implicit rejection value depends on the ciphertext and some secret value $s$ that is a part of the secret key.

\begin{figure}[H]
    \center
    \begin{minipage}{0.3\textwidth}
        \begin{algorithm}[H]
            \caption{$\keygen^{\not\bot}$}

        \end{algorithm}
    \end{minipage}
    \hfill
    \begin{minipage}{0.29\textwidth}

    \end{minipage}
    \hfill
    \begin{minipage}{0.3\textwidth}

    \end{minipage}
    \caption{$U^{\not\bot}$ routines}
\end{figure}
